\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-05-07 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add world things

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%\documentclass[acmsmall]{acmart}\settopmatter{}
%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,screen]{acmart}\settopmatter{}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10002986.10002989</concept_id>
<concept_desc>Security and privacy~Formal security models</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10002978.10003006</concept_id>
<concept_desc>Security and privacy~Systems security</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010124.10010138</concept_id>
<concept_desc>Theory of computation~Program reasoning</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{Security and privacy~Formal security models}
\ccsdesc[300]{Security and privacy~Systems security}
\ccsdesc[300]{Theory of computation~Program reasoning}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
% \acmJournal{PACMPL}
% \acmVolume{1}
% \acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
% \acmArticle{1}
% \acmYear{2018}
% \acmMonth{1}
% \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
% \startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
% \setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear
\usepackage[utf8]{inputenc}

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

% %% Some recommended packages.
% \usepackage{booktabs}   %% For formal tables:
%                         %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{wrapfig}

\usepackage{placeins}

\usepackage[disable]{todonotes}

\input{preamble-illu}

\input{../preamble}
% Paper specific redefinitions of commands
\renewcommand{\npair}[2][n]{#2}
\renewcommand{\MemSeg}{\shareddom{MemSeg}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}
\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
\renewcommand{\comp}{\var{comp}}
\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}

\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
\renewcommand{\lrr}[1][]{\lrrg[#1]{\untrusted}}
\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
\renewcommand{\lrv}[1][]{\lrvg[#1]{\untrusted}}
\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
\renewcommand{\lrp}{\typesetlr{P}}

\renewcommand{\lrrs}{\typesetlr{R}}
\renewcommand{\lrm}{\typesetlr{M}}
\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}

\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
\renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}

\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
\newcommand{\srcxjmpres}[1]{\plainfun{\srcalt{xjmpRes}}{#1}}
\newcommand{\wdjud}[2][ ]{#1 \vdash #2}

\newcommand{\trgcm}{\textsc{LCM}}
\newcommand{\srccm}{\textsc{oLCM}}
\newcommand{\extend}[1]{}

\setlength{\belowcaptionskip}{-5pt}

%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
%%%
%%% The following is specific to POPL '19 and the paper
%%% 'StkTokens: Enforcing Well-Bracketed Control Flow and Stack Encapsulation using Linear Capabilities'
%%% by Lau Skorstengaard, Dominique Devriese, and Lars Birkedal.
%%%
\setcopyright{rightsretained}
\copyrightyear{2019}
\acmJournal{PACMPL}
\acmYear{2019}
\acmVolume{3}
\acmNumber{POPL}
\acmArticle{19}
\acmMonth{1}
\acmPrice{}
\acmDOI{10.1145/3290332}

\begin{document}

%% Title information
\title[\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
% \titlenote{with title note}
% \subtitle{Fully abstract overlay semantics}
% \subtitlenote{with subtitle note}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

\author{Lau Skorstengaard}
% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \institution{Aarhus~University}
  \country{Denmark}                    %% \country is recommended
}
\email{lask@cs.au.dk}

%% Author with two affiliations and emails.
\author{Dominique Devriese}
\orcid{0000-0002-3862-6856}             %% \orcid is optional
\affiliation{
  \institution{Vrije~Universiteit~Brussel}           %% \institution is required
  \country{Belgium}                   %% \country is recommended
}
\email{dominique.devriese@vub.be}         %% \email is recommended

\author{Lars Birkedal}
% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \institution{Aarhus~University}
  \country{Denmark}                    %% \country is recommended
}
\email{birkedal@cs.au.dk}

\begin{abstract}
  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalize and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
\end{abstract}

% %% 2012 ACM Computing Classification System (CSS) concepts
% %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
% \begin{CCSXML}
% <ccs2012>
% <concept>
% <concept_id>10011007.10011006.10011008</concept_id>
% <concept_desc>Software and its engineering~General programming languages</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
% <concept>
% <concept_id>10003456.10003457.10003521.10003525</concept_id>
% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
% <concept_significance>300</concept_significance>
% </concept>
% </ccs2012>
% \end{CCSXML}

% \ccsdesc[500]{Software and its engineering~General programming languages}
% \ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}

\maketitle

\section{Introduction}
\label{sec:introduction}
\todo[inline]{TODOS ENABLED!!!}
\todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
Two properties that hold in many high-level source languages, but not in the assembly languages they are compiled to, are well-bracketed control flow and encapsulation of local state.

Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations. 
At the assembly level, this property does not hold immediately. 
Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
There is, however, no guarantee that untrusted assembly code respects this intended usage.
In particular, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.

Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
At the assembly level, this property also does not hold immediately.
The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.

To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
Capabilities come in different flavours.
Memory capabilities allow reading from and writing to a block of memory.
Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e. a piece of code coupled with private state that it gains access to upon invocation.
The concrete mechanics of object capabilities vary between different capability machines.
% KJAA: This 'sealed' and 'common seal' business is unclear to me.
For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
Each of the two capabilities are sealed with a common seal which make them opaque.
The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.

To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
The first is to use separate stacks for mutually distrusting components, and a central, trusted stack manager that mediates cross-component invocations.
This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
First, it requires reserving separate stack space for all components, which scales poorly to large amounts of components.
Also, in the presence of higher-order values (e.g., function pointers, objects), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
It is not clear how it can do this efficiently in the presence of large amounts of components.
Finally, this approach does not allow passing stack references between components.

A more scalable approach retains a single stack shared between components.
Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
\citet{skorstengaard_reasoning_2017} have demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
In fact, a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
% KJAA: 'Boundary crossing' is new here.
However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.

A second contribution of this work is the way in which we formulate these two properties.
Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.

>>>>

%\todo[inline]{Write small introduction. Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
% \begin{itemize}
% \item Logical relation
% \item FTLR
% \item Sketch high-level structure of the proof
% \end{itemize}
To prove Theorem~\ref{thm:full-abstraction}, we will essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates if and only if the other does.
The difficult part is to define when components are related.
In the next section, we give an informal overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.

\subsection{Kripke worlds}
\label{subsec:worlds}
The relation between \srccm{} and \trgcm{} components is non-trivial: essentially, we will say that components are related if invoking them with related values produces related observable behavior.
However, values are often only related under certain assumptions about the rest of the system.
For example, the linear data part of a return capability should only be related to the corresponding \srccm{} capability if no other value in the system references the same inactive stack frame and it is sealed with a seal only used for return pointers to the same code location.
To accommodate such conditional relatedness, we construct the relation as a step-indexed Kripke logical relation with recursive worlds.
\begin{jversion}
In this presentation, we try to include most of the details.
We refer to the conference version of this paper for a presentation\lau{insert reference} with fewer details and emphasis on the intuition.
\end{jversion}

\begin{jversion}
Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
A world is a semantic model of the memory.
In its simplest form, it is a collection of invariants that the memory must satisfy.
The invariants of a world can vary in complexity and expressiveness depending on the application.
The possible contents of the memory also influences what the world looks like.
For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} are modelled by the worlds.
In order to related \trgcm{} and \srccm{}, we need to model all the features of \srccm{} in the worlds which means we have to model:
\begin{itemize}
\item Three kinds of memory: heap, stack of local frames, and free stack
\item Linearity
\item Call stack
\item Seals
\end{itemize}

\subsubsection{Triple world and regions}
\srccm{}s memory is split in three: heap, free stack and encapsulated local stack frames.
In order to model the three kinds of memory, we simply have three sub-worlds.
That is, our world is defined as
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]
The sub-worlds are partial maps from names $\RegName$, modelled as natural numbers, to invariants.
In order to define what this actually means, we need to be more precise about what we mean by invariants.
We call the invariants regions because, as we will see later, they turn out to not be invariant.
A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments ($\Rel{\MemSeg \times \MemSeg}$).
Intuitively, we want to be able to say that two memory segments are related when their contents are related.
That is, for every address in the two memory segments, the words that resides there must be related.
In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now consider the intuition.
Say we have two integers, then they are related if they are equal.
If we instead have two capabilities, then they should also be related if they in some sense are equal, but what does it mean for capabilities to be ``equal''?
Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably do the same computation.
As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
To say whether two capabilities are related, we must know the possible contents of the memory.
This is exactly what our world express, so our regions will be world indexed, i.e.
\[
  \Worldh = \RegName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
\]
At this point, we can see that we have constructed a recursive domain equation.
If we inline $\Worldh$ in $\World$, then we have a circular equation with no solution because the self-reference happens in a negative position.
Luckily, we can solve circular equations if we move to a different domain.
For now, we will ignore the problem and return to the issue in Section~\ref{subsubsec:rec-dom-eq}.

For the sake of readability, we introduce the following notation
\begin{gather*}
  \pwheap = \pi_1(W)\\
  \pwpriv = \pi_2(W)\\
  \pwfree = \pi_3(W)
\end{gather*}

If we just wanted to model the memory, then the above regions would suffice, but we also need the world to model linearity and seals.
In the following subsections, we will gradually introduce the necessary components for our worlds.

\subsubsection{Linearity and world joins}
% Linearity
The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they can reference.
In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one capability references the unique parts.
%% Bookkeeping done by worlds
We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
If a memory segment is governed by a shared region, then non-linear capabilities may reference it.
On the other hand, if a memory segment is governed by a spatial region, then only linear capability may reference it.
%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
It is only the spatial owned regions that can be referenced by a capability.
Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})\uplus \\ 
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

% What parts of memory are linear
We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we have introduced enough to finish the definition of each of the three sub-worlds.
The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and non-linear capabilities, so it should contain both shared and spatial regions.
For this reason, it is defined as
\[
  \Worldh = \RegName \parfun (\World \fun \Regionh + \Regions)
\]
\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
For instance the $\Worldfs$ is defined as
\[
  \Worldfs = \RegName \parfun (\World \fun \Regions)
\]
$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
Conceptually, this means that each of the stack frames is connected with a return point in some code.
In a traditional C calling convention, the code return point would even be stored in the stack frame.
However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together constitute the sealed return pair.
In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
To this end, each shared region in $\Worlds$ must be paired with a return address:
\[
\Worlds = \RegName \parfun (\World \fun \Regions) \times \Addr 
\]

% Ensuring linearity
The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities uniquely references part of memory.

\subsubsection{Seals}
\label{subsubsec:seals}
% StkTokens designate seals for certain purposes, return, closure, etc
In order to guarantee well-bracketed control flow and local-state encapsulation, \stktokens{} designates seals for return capabilities and closure seals and relies on designated seals only being used for their intended purpose.
The designation of return seal is a system invariant that is modelled in the world as a \textit{seal interpretation function}.
This function takes a seal and returns a world-indexed relation that relates all the sealables that may be sealed with this specific seals.
That is
\[
  \Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps}
\]
%% Shared region has a seal indexed invariant that specifies what sealables can be sealed with a specific seal
In Section~\ref{subsubsec:ft-and-revocation}, we will see that not all regions are invariant.
The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
As we will see, shared regions are invariant, so we add the seal interpretation function as follows
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\World \fun \Rel{\MemSeg\times\MemSeg}) \times \\
  (\Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps})
\end{multline*}

\lau{Maybe add a bit more of explanation? Invariant needed for the LR to make sure that designated seals are used for their purpose. }

\subsubsection{Future worlds and revocation}
\label{subsubsec:ft-and-revocation}
% Memory evolves over time, we need to model this.
The world specifies the contents of the memory.
When a machine executes, the contents of the memory changes and parts of memory may be repurposed.
The world must model these changes.
% Future World
To this end, we have a future world relation $\future$ that describes how worlds may change over time.
For our purpose, relating \trgcm{} to \srccm{}, we are mainly interested in changes to the stack as it is the part that will be repurposed.
Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions that govern call frames that have been popped from the call stack, so the stack frame can be reused for a new call frame.
We allow introduction of new regions simply by extension.
That is, a future world can have new regions that was not present in a past world.
Things are a bit more complicated when it comes to revocation.
\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).}
For technical reasons, we need the worlds to be extensional which means that when it comes to removing regions in future worlds, we cannot simply forget about them.
Instead of removing the regions, we keep them around and tag them as removed with a $\revoked$ tag.
This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.
It is not all regions that need to be able to be revoked.
Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities in some sense can be revoked.
A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory.
This means that it is safe to repurpose that part of memory.
On the other hand, non-linear capabilities may have an alias, so it is not safe to repurpose the memory a non-linear capability has authority over.
It is only spatial regions that specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
That is, we have
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
    \{\revoked \}
  \end{array} \right.
\]
The rules for the future region relation are displayed in Figure~\ref{fig:ft-reg-rel}.
% Revocation of spatial regions. 'revocation' of linear capabilities.
All regions are future regions of themselves, so regions are allowed to stay unchanged.
The $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
% Spatial -> spatialo ~ affine ?
\todo[inline]{Lau: explain why $\spatialo \future \spatial$ gives us affine capabilitites. This is claimed in the technical report, but I don't see how this is different from revoking $\spatial$ regions and adding a new region with the future world relation.}
\begin{figure}[htb]
  \centering
  \begin{mathpar}
  \inferrule{ }{ \revoked \future (\spatial,\_)}
  \and
  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
  \and
  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
\end{mathpar}
  \caption{Future region relation.}
  \label{fig:ft-reg-rel}
\end{figure}

With the future region relation in place, we define the future world relation as follows:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
            { W' \future W }
\end{mathpar}
The relation says that each of the three worlds must be an extension of the past world and for each of the existing regions that must be a future region.
% injective function
Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard:esop18} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.

% TODO monotonicity of regions

\subsubsection{Solving the recursive domain equation}
\label{subsubsec:rec-dom-eq}
%solving the recursive domain equation means that we can present the actual definitions
In the previous sections, we have sketched what we would like our worlds to look like.
However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
% Move to a different domain
Therefore, we need to move to a different domain with enough structure for such an equation to have a solution.
Specifically, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
Intuitively, a c.o.f.e. is a set with some extra structure.
Specifically, c.o.f.e.'s have a step-index which provide sufficient structure for us to solve recursive domain equations.
In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.

% Cofe's
In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
We loosely follow the presentation in \citet{Birkedal:tutorial-notes}.
\begin{definition}[Ordered family of equivalences (o.f.e.)]
An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and for each $n$ $\nequal$ is an equivalence relation.
The pair must satisfy the following properties
\begin{enumerate}
\item $\nequal[0]$ is the total relation.
\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
\end{enumerate}
\end{definition}
\noindent One can think of the indexes on the family of equivalences as a measure of precision.
The larger $n$ is, the more refined the equivalence may be.
On the other hand as $n$ decreases, the equivalence becomes more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
At index 0, all precision is lost and the equivalence cannot distinguish anything.
% TODO: Lau write intuition for sequences and why they are necessary.
\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
  \label{def:cauchy-sequence}
  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
  in $X$.

  Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
  elements of $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
  \end{align*}
  or in words, the elements of the chain get arbitrarily close.

  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
  \end{align*}
\end{definition}
Functions between c.o.f.e.'s must retain the added structure relatively to sets.
To this end, we require all our functions to be \emph{non-expansive} which means that $n$-equivalences are preserved.
A function is called \emph{contractive} when it not only retain equivalences but makes them more precise.
\begin{definition}
  \label{def:nonexpansive-contractive-ofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
  We 
  The function $f$ is
  \begin{description}
  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal_Y f(x')
\]
  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
  \qedhere
\]
  \end{description}
\end{definition}
\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
However, this is not the only structure we would like to impose on our sets.
In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
The future world relations is essentially a preorder structure that we would like to impose on our sets.
We impose this structure on our sets by moving to \emph{preordered c.o.f.e.'s}.
% preordered cofe
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The new structure we add to the set, the preorder, must be consistent with the existing structure which is why it must preserve limits.
